# Tuango - RFM Analysis for Mobile App Push Messaging

* Name: Vidya Chockalingam
* Gitlab id: vichocka 
* Gitlab username: rsm-vichocka

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Loading data

```{r}
## make all required libraries available
suppressPackageStartupMessages(library(radiant))

## finding the path to the class dropbox folder
fp <- file.path(find_dropbox(),"MGTA455-2017")

## loading the data
## DO NOT change the next line!
## Data must be loaded from Dropbox/MGTA455-2017/data
loadr(file.path(fp, "data/tuango.rds"))
```

## Question answers

### Question 1

The percentage of buyers who responded to the promotions is 3.09%.
```{r}
#1
response_rate <- sum(tuango$buyer == "yes")/nrow(tuango)
response_rate
```

### Question 2
The average amount spent by customers who got at least one order is 202.356 RMB
```{r}
#2
avg_amnt_spent <- sum(tuango$ordersize)/sum(tuango$buyer == "yes")
avg_amnt_spent
```

### Question 3 & 4

As expected, the highest response rate is from quintile for all three - recency, frequency and monetary.
```{r}
#3 and 4
tuango$rec_iq <-  xtile(tuango$recency, 5)
visualize(dataset = "tuango", xvar = "rec_iq", yvar = "buyer", type = "bar", custom = TRUE) +  ylab("Proportion of buyer = 'yes'") + xlab("Recency quintiles (rec_iq)") + theme(legend.position = "none")

tuango$freq_iq <-  xtile(tuango$frequency, 5, rev=TRUE)
visualize(dataset = "tuango", xvar = "freq_iq", yvar = "buyer", type = "bar", custom = TRUE) +  ylab("Proportion of buyer = 'yes'") + xlab("Frequency quintiles (freq_iq)") + theme(legend.position = "none")

tuango$mon_iq <-  xtile(tuango$monetary, 5, rev=TRUE)
visualize(dataset = "tuango", xvar = "mon_iq", yvar = "buyer", type = "bar", custom = TRUE) +  ylab("Proportion of buyer = 'yes'") + xlab("Monetary quintiles (mon_iq)") + theme(legend.position = "none")

```
### Question 5
Please find below the plots for tha average amount spent by the buyers who responded to the promotions by quintiles.
```{r}
#5
tuango_responded <- tuango[tuango$buyer=="yes",]
by_recency <- group_by(tuango_responded,rec_iq)
avg_amnt_recency <- summarize(by_recency,avg_amnt_recency_quantile=mean(ordersize))
visualize(dataset = "avg_amnt_recency", xvar = "rec_iq", yvar = "avg_amnt_recency_quantile", type = "bar", custom = TRUE) +  ylab("Amount spent by proportion of buyer = 'yes'") + xlab("Recency quintiles (rec_iq)") + theme(legend.position = "none")


by_frequency <- group_by(tuango_responded,freq_iq)
avg_amnt_frequency <- summarize(by_frequency,avg_amnt_frequency_quantile=mean(ordersize))
visualize(dataset = "avg_amnt_frequency", xvar = "freq_iq", yvar = "avg_amnt_frequency_quantile", type = "bar", custom = TRUE) +  ylab("Amount spent by proportion of buyer = 'yes'") + xlab("Frequency quintiles (freq_iq)") + theme(legend.position = "none")

by_monetary <- group_by(tuango_responded,mon_iq)
avg_amnt_monetary <- summarize(by_monetary,avg_amnt_monetary_quantile=mean(ordersize))
visualize(dataset = "avg_amnt_monetary", xvar = "mon_iq", yvar = "avg_amnt_monetary_quantile", type = "bar", custom = TRUE) +  ylab("Amount spent by proportion of buyer = 'yes'") + xlab("Monetary quintiles (mon_iq)") + theme(legend.position = "none")
```

### Question 6
While the highest response rate is from quintile 1 for all three -frequency, recency and monetary, the average order size varies highly. The average order size/amount spent is not the highest for quintile 1 in all the three RFM.

### Question 7

Creating RFM indices using Independent quintile approach.
```{r}
#7
tuango$rfm_iq <- paste0(tuango$rec_iq,tuango$freq_iq,tuango$mon_iq)
```
Checking if rfm_iq matches with the refm_iq_pre provided in the dataset.
```{r}
#check
sum(tuango$rfm_iq == tuango$rfm_iq_pre)/nrow(tuango)
```

Creating RFM indices using Sequential quintile approach.
```{r}
tuango_sq <- tuango
tuango_sq <- group_by(tuango_sq, rec_iq) %>% mutate(freq_sq = xtile(frequency, 5, rev = TRUE)) %>% ungroup

## create new variable(s)
tuango_sq <- group_by(tuango_sq, rec_iq, freq_sq) %>% mutate(mon_sq = xtile(monetary, 5, rev = TRUE)) %>% ungroup
tuango_sq$rfm_sq <- paste0(tuango$rec_iq,tuango_sq$freq_sq,tuango_sq$mon_sq)

```


### Question 7
Break even response rate is the cost of sending emails/revenue Revenue is the average revenue from each deal - fee given to tuango. Fee given to tuango is 0.5* average amount spent by the customer.
```{r}
# Break-even rate
cost <- 1.6
Revenue <- 0.5*avg_amnt_spent
break_even_resp_rate <- cost/Revenue
break_even_resp_rate
```

### Question 8
If we give the deals to all buyers, then the expected number of buyers who will respond is response rate * the total number of people to whom the deals are sent.
Profit is the revenue from this response rate of people - the cost of sending the promotions to everyone  = 406725.4

Return on marketing is the profit/cost = 0.959
```{r}
#8
expected_no_buyers <- 264841*response_rate
profit <- expected_no_buyers*Revenue - cost*264841
profit
return_on_marketing <- profit/(cost*264841)
return_on_marketing
```


### Question 9 and 10
If we send the deals only to the people in cells which have a response rate greater than the break even rate, we can get a higher return on marketing (2.11) and a higher profit (520427).
```{r}
#9 & #10
tuango_sq <- group_by(tuango_sq, rfm_sq) %>% mutate(mailto_sq = mean(buyer == "yes") > break_even_resp_rate) %>% ungroup
perc_mail = mean(tuango_sq$mailto_sq)
nr_mail = 264841 * perc_mail
rep_rate = filter(tuango_sq, mailto_sq == TRUE) %>% summarize(mean(buyer == "yes")) %>% unlist
nr_resp = nr_mail * rep_rate
mail_cost = cost * nr_mail
profit_sq = (Revenue) * nr_resp - mail_cost
profit_sq
ROME_sq = profit_sq / mail_cost
ROME_sq
```


### Question 11
Since the average amount spent by customers belonging to each cell on the deals will be different across cells, it makes more sense to calculate a separate break even rate for each cell and see if the actual response rate of that cee is greater than the breakeven rate for that cell. 

This results in an even higher profit (595343) and return on marketing (2.15).
```{r}
#11
tuango_sq <- group_by(tuango_sq, rfm_sq) %>% mutate(breakevenrate_per_quint = 0.5*sum(ordersize)/sum(buyer == "yes")) %>% ungroup
tuango_sq$breakevenrate_per_quint <- 1.6/tuango_sq$breakevenrate_per_quint
tuango_sq <- group_by(tuango_sq, rfm_sq) %>% mutate(mailto_sq = mean(buyer == "yes") > breakevenrate_per_quint) %>% ungroup
perc_mail = mean(tuango_sq$mailto_sq,na.rm = TRUE)
nr_mail = 264841 * perc_mail
rep_rate = filter(tuango_sq, mailto_sq == TRUE) %>% summarize(mean(buyer == "yes")) %>% unlist
nr_resp = nr_mail * rep_rate
mail_cost = cost * nr_mail
profit_sq = (Revenue) * nr_resp - mail_cost
profit_sq
ROME_sq = profit_sq / mail_cost
ROME_sq
```

### Question 12
Since there will be an error in our response rate estimation, we need to calulate a 95% confidence interval for the response rate and then compare it to the breakeven rate. This leads to fewer cells being targeted and hence the cost reduces drastically.
This leads to almost double the return on marketing but the profits are lower.
```{r}
#12
tuango_sq <- group_by(tuango_sq, rfm_sq) %>% mutate(mailto_csq = (mean(buyer == "yes") - 1.96 * sqrt(mean(buyer == "yes")*(1-mean(buyer=="yes"))/n()))  > breakevenrate_per_quint) %>% ungroup
perc_mail = mean(tuango_sq$mailto_csq,na.rm = TRUE)
nr_mail = 264841 * perc_mail
rep_rate = filter(tuango_sq, mailto_csq == TRUE) %>% summarize(mean(buyer == "yes")) %>% unlist
nr_resp = nr_mail * rep_rate
mail_cost = cost * nr_mail
profit_sq = (Revenue) * nr_resp - mail_cost
profit_sq
ROME_sq = profit_sq / mail_cost
ROME_sq
```
