# BBB - Logistic Regression

* Name: Vidya Chockalingam
* Gitlab id: rsm-vichocka
* Gitlab username: vichocka

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
## make all required libraries available
suppressPackageStartupMessages(library(radiant))
library(dplyr)
#install.packages("caret")
library(caret)
## finding the path to the class dropbox folder
fp <- file.path(find_dropbox(),"MGTA455-2017")

## loading the data
## DO NOT change the next line!
## Data must be loaded from Dropbox/MGTA455-2017/data
loadr(file.path(fp, "data/bbb.rda"))
```

## Question answers
### Part 1
Building a logistic regression model with buyer as the dependent variable and gender, last, total etc. as the independent variables.
```{r}
#Part 1
#1
bbb$buyer_int <- ifelse(bbb$buyer == "yes",1 ,0) 
response_rate <- sum(bbb$buyer_int)/nrow(bbb)
model <- glm(buyer_int ~ gender + last + total + child + youth + cook + do_it +reference + art + geog,family=binomial(link='logit'),data=bbb)
bbb$purch_prob <- predict(model, newdata = bbb, type = "response")
```

All the p values are <0.05 and hence with a 95% confidence interval we can reject the null hypothesis of having the coefficients of the independent variables to be 0.
```{r}
#2
summary(model)
```

### Part 2
Splitting the data into deciles based on the predicted scores of the logit model.
```{r}
# Part 2
#1
bbb$prob_dec <-  xtile(bbb$purch_prob, 10, rev=TRUE)
by_decile <- group_by(bbb,prob_dec)
avg_score <- summarize(by_decile,avg_score=mean(purch_prob))
```

Plotting the response rate across deciles - making sure that the highest response rate is in the first decile.
```{r}
#2
visualize(dataset = "bbb", xvar = "prob_dec", yvar = "buyer_int", type = "bar", custom = TRUE) +  ylab("Proportion of buyer = 'yes'") + xlab("Deciles (freq_iq)") + theme(legend.position = "none")
```

Storing the response rate across deciles in a table.
```{r}
#3
no.of.customers <- summarize(by_decile,customer_count=n())
no.of.customers.bought <- summarize(by_decile,customer_count_yes=sum(buyer_int))
response.rate <- summarize(by_decile,response_rate=sum(buyer_int)/n())
```

Creating another glm model with child as the only independent variable.
```{r}
#4
model1 <- glm(buyer_int ~ child ,family=binomial(link='logit'),data=bbb)
summary(model1)
```

Creating Lift and Gain charts for the logit model.
```{r}
# Part 3
#1
lift_table <- cbind(no.of.customers,no.of.customers.bought,response.rate)
lift_table <- lift_table[,c(-3,-5)]
for(i in 1:nrow(lift_table))
{
  if(i==1)
    {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")]
  }
  else
  {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")] + lift_table$cum_cust_count[i-1]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")] + lift_table$cum_buyer_count[i-1]
  }
}
lift_table$cum_prop <- lift_table$cum_cust_count/nrow(bbb)
lift_table$lift <- lift_table$response_rate/response_rate
lift_table$cum_resp_rate <- lift_table$cum_buyer_count/lift_table$cum_cust_count
lift_table$cum_lift <- lift_table$cum_resp_rate/response_rate
#2
visualize(dataset = "lift_table", xvar = "cum_prop", yvar = "cum_lift", type = "line", custom = TRUE) +  ylab("Cumulative Lift") + xlab("Proportion of customers") + theme(legend.position = "none")
```
Gain Chart
```{r}
#3
gain_table <- lift_table
gain_table$buyer_prop <- gain_table$customer_count_yes/sum(gain_table$customer_count_yes)
for(i in 1:nrow(gain_table))
{
  if(i==1)
    {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")]
  }
  else
  {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")] + gain_table$cum_buyer_prop[i-1]
  }
}
#4
visualize(dataset = "gain_table", xvar = "cum_prop", yvar = "cum_buyer_prop", type = "line", custom = TRUE) +   geom_segment(aes(x = min(cum_prop), y = min(cum_buyer_prop), xend = max(cum_prop), yend = max(cum_buyer_prop), colour = "segment"), data = gain_table) + ylab("Cumulative Gain") + xlab("Proportion of customers") + theme(legend.position = "none")
```

### Part 4
Creating the Profit and ROME curve using the predictions by the logit model.
```{r}
#Part 4
mail <- 0.50
price <- 9
shipping <- 3
selling_price <- 18
#1
lift_table$profit <- (selling_price-price-shipping)*lift_table$customer_count_yes - mail*lift_table$customer_count
lift_table$ROME <- lift_table$profit/(mail*lift_table$customer_count)
#2
ggplot(lift_table,aes(x=prob_dec,y=profit,group=1)) + geom_line() 
#3
ggplot(lift_table,aes(x=prob_dec,y=ROME,group=1)) + geom_line() 
```

Creating the confusion matrix for the above model.
```{r}
#4
bbb$prediction <- ifelse(bbb$purch_prob>=0.5, 1, 0)
confusion_table <- table(bbb$prediction,bbb$buyer_int)
print(confusion_table)
confusion_matrix <- confusionMatrix(confusion_table)
confusion_matrix$overall['Accuracy']
```

### Part 5
Calculating the expected number of buyers, reponse rate of the responders, profit and ROME if we target only the customers who have a purchase probability > break even rate.
```{r}
#Part 5
#1
break_even_rate <- mail/(selling_price - shipping - price)
#2
bbb$mailto_logit <- ifelse(bbb$purch_prob>break_even_rate,1,0)
#3
no.of.customers.to.target <- mean(bbb$mailto_logit)
no.of.customers <- 500000 * no.of.customers.to.target
print(paste("Targeted Customers ",no.of.customers))
bbb_responders <- bbb[bbb$mailto_logit == 1,]
response_rate_responders <- sum(bbb_responders$buyer_int)/nrow(bbb_responders)
print(paste("Response Rate of the Responders ",response_rate_responders))
expected.buyers <- round(no.of.customers * response_rate_responders)
print(paste("Expected number of buyers ",expected.buyers))
#4
profit <- (selling_price-shipping-price)*expected.buyers - mail*no.of.customers
print(paste("Profit if we target only the customers with purchase probability > break even rate is $",profit))
ROME_targeted <- profit/(mail*no.of.customers)
ROME_targeted
```
## Logistic Regression Accounting for Standard Error

If we predict logistic regression accounting for standard error in the prediction of the parameters, then we find that there is not much variation in the ROME, accuracy and profit.
```{r}
#5 Logistic with Error
loadr(file.path(fp, "data/bbb.rda"))
bbb_ler <- bbb
bbb_ler$buyer_int <- ifelse(bbb_ler$buyer == "yes",1 ,0) 
bbb_ler$purch_prob <- predict(model, newdata = bbb_ler, type = "response",se = TRUE)$fit - predict(model, newdata = bbb_ler, type = "response",se = TRUE)$se.fit
bbb_ler$prob_dec <-  xtile(bbb_ler$purch_prob, 10, rev=TRUE)

by_decile <- group_by(bbb_ler,prob_dec)
no.of.customers <- summarize(by_decile,customer_count=n())
no.of.customers.bought <- summarize(by_decile,customer_count_yes=sum(buyer_int))
response.rate <- summarize(by_decile,response_rate=sum(buyer_int)/n())
```

Lift and Gains Table for Logistic regression with standard error.
```{r}
# Part 3
#1
lift_table <- cbind(no.of.customers,no.of.customers.bought,response.rate)
lift_table <- lift_table[,c(-3,-5)]
for(i in 1:nrow(lift_table))
{
  if(i==1)
    {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")]
  }
  else
  {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")] + lift_table$cum_cust_count[i-1]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")] + lift_table$cum_buyer_count[i-1]
  }
}
lift_table$cum_prop <- lift_table$cum_cust_count/nrow(bbb)
lift_table$lift <- lift_table$response_rate/response_rate
lift_table$cum_resp_rate <- lift_table$cum_buyer_count/lift_table$cum_cust_count
lift_table$cum_lift <- lift_table$cum_resp_rate/response_rate
#2
visualize(dataset = "lift_table", xvar = "cum_prop", yvar = "cum_lift", type = "line", custom = TRUE) +  ylab("Cumulative Lift") + xlab("Proportion of customers") + theme(legend.position = "none")

#3
gain_table <- lift_table
gain_table$buyer_prop <- gain_table$customer_count_yes/sum(gain_table$customer_count_yes)
for(i in 1:nrow(gain_table))
{
  if(i==1)
    {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")]
  }
  else
  {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")] + gain_table$cum_buyer_prop[i-1]
  }
}
#4
visualize(dataset = "gain_table", xvar = "cum_prop", yvar = "cum_buyer_prop", type = "line", custom = TRUE) +   geom_segment(aes(x = min(cum_prop), y = min(cum_buyer_prop), xend = max(cum_prop), yend = max(cum_buyer_prop), colour = "segment"), data = gain_table) + ylab("Cumulative Gain") + xlab("Proportion of customers") + theme(legend.position = "none")
```

Profit and ROME curve for Logistic Regression with Standard Error
```{r}
#Part 4
#1
lift_table$profit <- (selling_price-price-shipping)*lift_table$customer_count_yes - mail*lift_table$customer_count
lift_table$ROME <- lift_table$profit/(mail*lift_table$customer_count)
#2
ggplot(lift_table,aes(x=prob_dec,y=profit,group=1)) + geom_line() 
#3
ggplot(lift_table,aes(x=prob_dec,y=ROME,group=1)) + geom_line() 
```

Accuracy for this model.
```{r}
#4
bbb_ler$prediction <- ifelse(bbb_ler$purch_prob>=0.5, 1, 0)
confusion_table <- table(bbb_ler$prediction,bbb_ler$buyer_int)
print(confusion_table)
confusion_matrix <- confusionMatrix(confusion_table)
confusion_matrix$overall['Accuracy']
```

```{r}
#2
bbb_ler$mailto_logit <- ifelse(bbb_ler$purch_prob>break_even_rate,1,0)
#3
no.of.customers.to.target <- mean(bbb_ler$mailto_logit)
no.of.customers <- 500000 * no.of.customers.to.target
print(paste("Number of customers to target", no.of.customers))
bbb_responders <- bbb_ler[bbb_ler$mailto_logit == 1,]
response_rate_responders <- sum(bbb_responders$buyer_int)/nrow(bbb_responders)
print(paste("Response rate expected",response_rate_responders))
expected.buyers <- round(no.of.customers * response_rate_responders)
print(paste("Expected Number of Buyers",expected.buyers))

#4
profit <- (selling_price-shipping-price)*expected.buyers - mail*no.of.customers
ROME_targeted <- profit/(mail*no.of.customers)
print(paste("Profit if we target only the customers with purchase probability > break even rate is $",profit))
print(paste("ROME for these targeted customers",ROME_targeted))

```

## Sequential RFM

Creating quintiles based on the recency(last), frequency(purch), monetary(total) values of the dataset.
```{r}
# Sequential RFM
bbb_sq <- bbb
bbb_sq$buyer_int <- ifelse(bbb$buyer == "yes",1 ,0) 
response_rate <- sum(bbb_sq$buyer_int)/nrow(bbb_sq)
bbb_sq$rec_iq <-  xtile(bbb_sq$last, 5)
bbb_sq <- group_by(bbb_sq, rec_iq) %>% mutate(freq_sq = xtile(purch, 5, rev = TRUE)) %>% ungroup

## create new variable(s)
bbb_sq <- group_by(bbb_sq, rec_iq, freq_sq) %>% mutate(mon_sq = xtile(total, 5, rev = TRUE)) %>% ungroup
bbb_sq$rfm_sq <- paste0(bbb_sq$rec_iq,bbb_sq$freq_sq,bbb_sq$mon_sq)

```

Finding the profit and ROME of the customers in the cells which has a response rate > break even rate.
```{r}
bbb_sq <- group_by(bbb_sq, rfm_sq) %>% mutate(mailto_sq = mean(buyer == "yes") > break_even_rate) %>% ungroup
perc_mail = mean(bbb_sq$mailto_sq)
nr_mail = 500000 * perc_mail
rep_rate = filter(bbb_sq, mailto_sq == TRUE) %>% summarize(mean(buyer == "yes")) %>% unlist
nr_resp = nr_mail * rep_rate
mail_cost = mail * nr_mail
profit_sq = (selling_price-shipping-price) * nr_resp - mail_cost
print(paste("Profit if we target only the customers with purchase probability > break even rate is $",profit_sq))
ROME_sq = profit_sq / mail_cost
print(paste("ROME ",ROME_sq))
```

Creating Lift and Gain Charts for Sequential RFM based on the response rate per rfm cell.
```{r}
bbb_sq <- group_by(bbb_sq, rfm_sq) %>% mutate(response_rate = mean(buyer == "yes")) %>% ungroup %>% mutate(prob_dec = xtile(response_rate,10, rev=TRUE)) %>% ungroup
by_decile <- group_by(bbb_sq,prob_dec)
no.of.customers <- summarize(by_decile,customer_count=n())
no.of.customers.bought <- summarize(by_decile,customer_count_yes=sum(buyer_int))
response.rate <- summarize(by_decile,response_rate=sum(buyer_int)/n())
lift_table <- cbind(no.of.customers,no.of.customers.bought,response.rate)
lift_table <- lift_table[,c(-3,-5)]
for(i in 1:nrow(lift_table))
{
  if(i==1)
    {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")]
  }
  else
  {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")] + lift_table$cum_cust_count[i-1]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")] + lift_table$cum_buyer_count[i-1]
  }
}
lift_table$cum_prop <- lift_table$cum_cust_count/nrow(bbb)
lift_table$lift <- lift_table$response_rate/response_rate
lift_table$cum_resp_rate <- lift_table$cum_buyer_count/lift_table$cum_cust_count
lift_table$cum_lift <- lift_table$cum_resp_rate/response_rate
#2
visualize(dataset = "lift_table", xvar = "cum_prop", yvar = "cum_lift", type = "line", custom = TRUE) +  ylab("Cumulative Lift") + xlab("Proportion of customers") + theme(legend.position = "none")
#3
gain_table <- lift_table
gain_table$buyer_prop <- gain_table$customer_count_yes/sum(gain_table$customer_count_yes)
for(i in 1:nrow(gain_table))
{
  if(i==1)
    {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")]
  }
  else
  {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")] + gain_table$cum_buyer_prop[i-1]
  }
}
visualize(dataset = "gain_table", xvar = "cum_prop", yvar = "cum_buyer_prop", type = "line", custom = TRUE) +geom_segment(aes(x = min(cum_prop), y = min(cum_buyer_prop), xend = max(cum_prop), yend = max(cum_buyer_prop), colour = "segment"), data = gain_table) +  ylab("Cumulative Gain") + xlab("Proportion of customers") + theme(legend.position = "none")
```

Creating Profit and ROME curves for sequential RFM.
```{r}
lift_table$profit <- (selling_price-price-shipping)*lift_table$customer_count_yes - mail*lift_table$customer_count
lift_table$ROME <- lift_table$profit/(mail*lift_table$customer_count)
#2
ggplot(lift_table,aes(x=prob_dec,y=profit,group=1)) + geom_line() 
#3
ggplot(lift_table,aes(x=prob_dec,y=ROME,group=1)) + geom_line() 
```

Confusion Matrix and Accuracy for sequential RFM is calculated by taking the prediction scores for all the customers in a RFM cell with response rate > break even rate as 1. Then we compare it with the actual response rate of the customers in those cells.
```{r}
#Confusion matrix
bbb_sq$buyer_int <- ifelse(bbb_sq$buyer == "yes",1 ,0) 
bbb_sq$mailto_sq <- ifelse(bbb_sq$mailto_sq == 1, 1, 0)
confusion_table <- table(bbb_sq$mailto_sq,bbb_sq$buyer_int)
print(confusion_table)
confusion_matrix <- confusionMatrix(confusion_table)
confusion_matrix$overall['Accuracy']
```

## Naive Bayes
Calculating predictions based on Naive Bayes Model.
```{r}
## Naive Bayes
bbb_nb <- bbb
bbb_nb$buyer_int <- ifelse(bbb_nb$buyer == "yes",1 ,0) 
response_rate <- sum(bbb_nb$buyer_int)/nrow(bbb_nb)
levels(bbb_nb$buyer) <- c(0,1)
nb_model <- nb(dataset = "bbb_nb",rvar = "buyer", evar = c("gender","last","total","child","youth","cook","do_it", "reference", "art", "geog"))
bbb_nb$purch_prob <- predict(nb_model, pred_data  = "bbb_nb")[,"0"]
```

Plotting Lift and Gain Charts for Naive Bayes
```{r}
# Part 2
#1
bbb_nb$prob_dec <-  xtile(bbb_nb$purch_prob, 10, rev=TRUE)
by_decile <- group_by(bbb_nb,prob_dec)
no.of.customers <- summarize(by_decile,customer_count=n())
no.of.customers.bought <- summarize(by_decile,customer_count_yes=sum(buyer_int))
response.rate <- summarize(by_decile,response_rate=sum(buyer_int)/n())
lift_table <- cbind(no.of.customers,no.of.customers.bought,response.rate)
lift_table <- lift_table[,c(-3,-5)]
for(i in 1:nrow(lift_table))
{
  if(i==1)
    {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")]
  }
  else
  {
    lift_table$cum_cust_count[i] <- lift_table[i,c("customer_count")] + lift_table$cum_cust_count[i-1]
    lift_table$cum_buyer_count[i] <- lift_table[i,c("customer_count_yes")] + lift_table$cum_buyer_count[i-1]
  }
}
lift_table$cum_prop <- lift_table$cum_cust_count/nrow(bbb)
lift_table$lift <- lift_table$response_rate/response_rate
lift_table$cum_resp_rate <- lift_table$cum_buyer_count/lift_table$cum_cust_count
lift_table$cum_lift <- lift_table$cum_resp_rate/response_rate
#2
visualize(dataset = "lift_table", xvar = "cum_prop", yvar = "cum_lift", type = "line", custom = TRUE) +  ylab("Cumulative Lift") + xlab("Proportion of customers") + theme(legend.position = "none")

#3
gain_table <- lift_table
gain_table$buyer_prop <- gain_table$customer_count_yes/sum(gain_table$customer_count_yes)
for(i in 1:nrow(gain_table))
{
  if(i==1)
    {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")]
  }
  else
  {
    gain_table$cum_buyer_prop[i] <- gain_table[i,c("buyer_prop")] + gain_table$cum_buyer_prop[i-1]
  }
}
#4
visualize(dataset = "gain_table", xvar = "cum_prop", yvar = "cum_buyer_prop", type = "line", custom = TRUE) +   geom_segment(aes(x = min(cum_prop), y = min(cum_buyer_prop), xend = max(cum_prop), yend = max(cum_buyer_prop), colour = "segment"), data = gain_table) + ylab("Cumulative Gain") + xlab("Proportion of customers") + theme(legend.position = "none")
```

Plotting ROME and Profit curves across deciles in Naive Bayes.
```{r}
#1
lift_table$profit <- (selling_price-price-shipping)*lift_table$customer_count_yes - mail*lift_table$customer_count
lift_table$ROME <- lift_table$profit/(mail*lift_table$customer_count)
#2
ggplot(lift_table,aes(x=prob_dec,y=profit,group=1)) + geom_line() 
#3
ggplot(lift_table,aes(x=prob_dec,y=ROME,group=1)) + geom_line() 
```

Creating the confusion matrix and accuracy using Naive Bayes Model.
```{r}
#4
bbb_nb$prediction <- ifelse(bbb_nb$purch_prob>=0.5, 1, 0)
confusion_table <- table(bbb_nb$prediction,bbb_nb$buyer_int)
print(confusion_table)
confusion_matrix <- confusionMatrix(confusion_table)
confusion_matrix$overall['Accuracy']
```

Calculating the profit and ROME using Naive Bayes.
```{r}
bbb_nb$mailto_logit <- ifelse(bbb_nb$purch_prob>break_even_rate,1,0)
#3
no.of.customers.to.target <- mean(bbb_nb$mailto_logit)
no.of.customers <- 500000 * no.of.customers.to.target
bbb_responders <- bbb_nb[bbb_nb$mailto_logit == 1,]
response_rate_responders <- sum(bbb_responders$buyer_int)/nrow(bbb_responders)
expected.buyers <- round(no.of.customers * response_rate_responders)

#4
profit <- (selling_price-shipping-price)*expected.buyers - mail*no.of.customers
ROME_targeted <- profit/(mail*no.of.customers)
print(paste("Profit is $", profit))
ROME_targeted
print(paste("ROME", ROME_targeted))
```
